### 2019-03-16 README.Rmd
Update README
Use public CITES db release, bump to 0.1.0, +misc
- Change names in database
- Add convenience functions for metadata
- More completee README ready for public release
- Add metadata documentation from public DB
- Add SQL preview to connection pane

Features:
-  Local caching of versions of database
-  Metadata in **[dataspice](https://github.com/ropenscilabs/dataspice)**/<schema.org> format
-  RStudio connection pane with browsable database
-  HTMLwidget-enhanced help files for browsing metadata?

@@

```{r example}
#library(citesdb)

# some example code
```

# Current Verison

```{r version}
#packageVersion("citesdb")
```

## Contributing

@@

>>>>

**citesdb** is a package to conveniently analyze the full CITES shipment-level wildlife trade database, available at <https://trade.cites.org/>.

More details on the database can be found in the `?guidance` help file.

@@

### Getting the data

When you first load the package you will see a message like this:

    library(citesdb)
    #> Local CITES database empty or corrupt. Download with cites_db_download()

Not to worry, just do as it says and run `cites_db_download()`.  This will
fetch the most recent database from online, an approximately 158 MB download.
It will expand to over 1 GB in the local database. During
the downloaad and database building up to 3.5 GB of disk space may be used temporarily.

### Using the Database

Once you fetch the data you can connect to the database with the `cites_db()`
command.  You can use the
`cites_shipments()` command to load a remote `tibble` that is backed by the
database but not loaded into R.  You can use this
to analyze CITES data without ever loading it into memory, then gather your
results with `collect()`. For example:

```{r, warning = FALSE}
library(citesdb)
library(dplyr)
start <- Sys.time()

cites_shipments() %>%
  group_by(Year) %>%
  summarize(number = n()) %>%
  arrange(desc(Year)) %>%
  collect()

stop <- Sys.time()
```

The back-end database, [MonetDB](https://monetdb.org) is very fast and powerful, making such
analyses quite snappy even on such large data and normal desktops and laptops. Here's the
timing of the above query, which procecsses over 20 million records, on an ordinary laptop:

```{r}
stop - start
```

If you are using a recent version of RStudio interactively, loading the CITES package also brings
up a browsable pane in the "Connections" tab that lets you explore and preview
the database, as well as interact with it direcctly via SQL commands.

### Metadata

The database also contains tables of field metadata, codes used, and CITES countries. This information
comes from ["A guide to using the CITES Trade Database"](https://trade.cites.org/cites_trade_guidelines/en-CITES_Trade_Database_Guide.pdf),
on the CITES website.  Convenience functions `cites_metadata()`, `cites_codes()`, and `cites_parties()`
access this information:

```{r}
head(cites_metadata())

head(cites_codes())
```

More information on the release of shipment-level data can be found in the
`?guidance` help file.

### Contributing

@@

[![EcoHealth Allinace](https://github.com/ecohealthalliance/citesdb/blob/master/inst/img/eha-footer.png)](https://www.ecohealthalliance.org/)
